---
title: "SDS 192 Assignment"
subtitle: "SDS 192"
author: "Malaika Kironde, Jack Kendrick, Lika Mikhelashvili"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE,
  warning = FALSE
)
library(tidyverse)
library(sds192)
library(fivethirtyeight)
library(ggthemes)
```

## Introduction

<<<<<<< HEAD
In the first six months of 2015, [467 people in the US were killed by the police](https://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/). Using data from FiveThirtyEight, we explore the relationships between household income, race, and the number of people killed by the police. According to [US Census Bureau data](https://www.census.gov/library/publications/2016/demo/p60-256.html#:~:text=Median%20household%20income%20in%20the,before%20the%20most%20recent%20recession.), the national median household income in 2015 was $56.516, yet we see that the vast majority of people killed by the police across all races lived in neighborhoods with a median household income well below this national average.
=======
In the first six months of 2015, [467 people in the US were killed by the police](https://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/). Using data from FiveThirtyEight, we explore the relationships between household income, race, and the number of people killed by the police. According to [US Census Bureau data](https://www.census.gov/library/publications/2016/demo/p60-256.html#:~:text=Median%20household%20income%20in%20the,before%20the%20most%20recent%20recession.), the national median household income in 2015 was $56.516, yet we see that the vast majority of people killed by the police lived in neighborhoods with a median household income well below this national average. In fact, we see that the majority of people killed by the police across all races lived in neighborhoods where the median household income would place them in the lower and lower middle classes.
>>>>>>> fc67337064512a64ebe069080dc8faae2120d8df

We also examine the racial distribution of people killed by the police as it relates to geography. In almost every state and region of the US, we see that the number of white people killed by the police exceeds the number of people of color killed by the police. However, when we account for demographic, we find that the number of people of color killed by the police is disproportionately high.

```{r, include=FALSE}
# Adding in column for Region to dataset
state_region <- data.frame(state = state.abb, region = state.region)
state_region <- rbind(state_region, data.frame(state = "DC", region = "South"))
police_data <- merge(police_killings, state_region, by = "state")
```

```{r, include=FALSE}
police_data
```


### Graph 1

```{r graph1, message=FALSE}
#Plot Code - Wrapping by Race
ggplot(data=police_data, aes(h_income)) +
  geom_histogram(color="black",fill="white") +
  theme_fivethirtyeight()+ 
  geom_vline(xintercept=40000, linetype="dotted") +
  geom_text(aes(x=40000, label="Lower Class", y=20), color="blue", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=40000, label="Middle Class", y=20), color="blue", angle=90, vjust=1.5, size=2) +
  geom_text(aes(x=120000, label="Upper Class", y=20), color="blue", angle=90, vjust=1.5, size=2) +
  geom_vline(xintercept=120000, linetype="dotted") +
  theme(axis.title = element_text())+
  scale_fill_fivethirtyeight()+
  facet_wrap(~raceethnicity) + 
  scale_x_continuous(name = "Tract-level Median Household Income ($)")+
  scale_y_continuous(name = "Number of Individuals Killed")+
  labs(
    title = "Household Incomes Of People Killed by the Police", 
    subtitle = "Figure 1: Economic Breakdown of Police Killings",
    caption = "Source: fivethirtyeight"
  )
```

Graph 1 shows how the number of individuals killed varies with their median household income. Generally, we can see that for each race, the highest number of individuals killed by the police lies in lower class. The exception to this is Asian/Pacific Islander and White, where the highest number of individuals killed lies just above the boundary for the lower class. This is especially telling since even though the lower class only accounts for around 30% of the population (Pew Research Center, 2015), they still have the highest number of individuals killed by the police.

### Graph 2

```{r state_breakdown, message=FALSE}
ggplot(data = police_data, aes(state))+
  coord_flip()+
  geom_bar(aes(x = state, fill = raceethnicity))+
    labs(
    title = "Figure 2.1: Racial Breakdown of Police Killings", 
    subtitle = "Racial Distribution in each State",
    caption = "Source: FiveThirtyEight",
    fill = "Race/Ethnicity"
  )+
  scale_x_discrete(name = "State")+
  scale_y_continuous(name = "Number of Individuals Killed")+
  theme_fivethirtyeight()+ 
  theme(axis.title = element_text())+
  theme(axis.text.y = element_text(size  = 5))
```

```{r region_breakdown, message= FALSE}
ggplot(data = police_data, aes(region))+
  geom_bar(aes(x = region, fill = raceethnicity))+
  labs(title = "Figure 2.2: Racial Breakdown of Police Killings",
  subtitle = "Racial Distribution in each US Region",
  caption = "Source: FiveThirtyEight",
  fill = "Race/Ethnicity"
  )+
  scale_x_discrete(name = "Region")+
  scale_y_continuous(name = "Number of Individuals Killed")+
  theme_fivethirtyeight()+ 
  theme(axis.title = element_text())
```

<<<<<<< HEAD
The graphs show the racial breakdowns of the individuals killed by the police in 2015 in each state (Figure 2.1) and region (Figure 2.2). Each race is mapped to a specific color. In figure 2.1 it is evident that in most states, the majority of the killings were of White race. However, this data should not be taken out of context. We should expect that a higher number of white people are killed by the police because white people make up the majority of the US population. White people of non-Hispanic origin [account for around 60% of the population](https://www.census.gov/quickfacts/fact/table/US/PST045219), whereas Black people, the next most frequently killed demographic, account for only 13.4% of the population. Although we see higher numbers of white people being killed, when racial demographics are take into account, the number of people of color killed by the police is disproportionately high ^[Feldman, J., Gruskin, S., Coull, B., & Krieger, N. (2019). Police-Related Deaths and Neighborhood Economic and Racial/Ethnic Polarization, United States, 2015–2016. American Journal Of Public Health, 109(3), 458-464. https://doi.org/10.2105/ajph.2018.304851].
=======

The graphs show the racial breakdowns of the individuals killed by the police in 2015 in each state (Figure 2.1) and region (Figure 2.2). Each race is mapped to a specific color. In figure 2.1 it is evident that in most states, the majority of the killings were of White race. However, a large proportion of Black individuals were killed in almost all states. Such trend is more evident in figure 2.2, where among all minority groups (Asian/Pacific Islander, Black Hispanic/Latino, Native American), a drastically high number of Black individuals are killed in almost all regions. 
>>>>>>> fc67337064512a64ebe069080dc8faae2120d8df


## Word count

```{r word_count, message=FALSE, echo=FALSE}
text_stats()
```

## Standards

In this assignment, we attempted the following [standards](https://beanumber.github.io/sds192/standards.html):

- `r standard_button("markdown")`
We met the Markdown standard because we were able to format our markdown file such that it shows only necessary content by usinG booleans such as include, eval and echo. We were able to format our file to be readable using code chunks and data graphics. Furthermore, we had titles, subtitles and paragraphs to add partitions to our markdown file.
- `r standard_button("aesthetics")`
We met the Aesthetics standard because we used a histogram for our first data graphic and a bar chart for our second data graphic. We used a categorical color scheme to show different races in graph 2.
- `r standard_button("context")`
We met the Context standard because we created graphics with readable axis labels and appropriate units and included legends where appropriate.
- `r standard_button("ethics")`
We met the Ethics standard because we acknowledged the potential ethical issues of displaying the racial breakdowns of people killed by the police and explained that this data should not be taken out of context.

