---
title: "SDS 192 Assignment"
subtitle: "SDS 192"
author: "Malaika Kironde, Jack Kendrick, Lika Mikhelashvili"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE
)
library(tidyverse)
library(sds192)
library(fivethirtyeight)
library(ggthemes)
```

## Introduction

```{r echo=FALSE}
# Adding in column for Region to dataset
state_region <- data.frame(state = state.abb, region = state.region)
state_region <- rbind(state_region, data.frame(state = "DC", region = "South"))

police_data <- merge(police_killings, state_region, by = "state")
```

### Graph 1

```{r graph1, message=FALSE}
#Plot Code - Wrapping by Race
ggplot(data=police_data, aes(h_income)) +
  geom_histogram(color="black",fill="white") +
  theme_fivethirtyeight()+ 
  geom_vline(xintercept=40000, linetype="dotted") +
  geom_text(aes(x=40000, label="Lower Class", y=20), color="blue", angle=90, vjust = -0.5, size=2) + 
  geom_text(aes(x=40000, label="Middle Class", y=20), color="blue", angle=90, vjust=1.5, size=2) +
  geom_text(aes(x=120000, label="Upper Class", y=20), color="blue", angle=90, vjust=1.5, size=2) +
  geom_vline(xintercept=120000, linetype="dotted") +
  theme(axis.title = element_text())+
  scale_fill_fivethirtyeight()+
  facet_wrap(~raceethnicity) + 
  scale_x_continuous(name = "Tract-level Median Household Income")+
  scale_y_continuous(name = "Number of Individuals Killed")+
  labs(
    title = "Median Household Income Distribution Among People Killed by the Police in 2015", 
    subtitle = "Distribution by Race",
    caption = "Source: fivethirtyeight"
  )
```

Graph 1 shows how the number of individuals killed varies with their median household income. Generally, we can see that for each race, the highest number of individuals killed by the police lies in lower class. The exception to this is Asian/Pacific Islander and White, where the highest number of individuals killed lies just above the boundary for the lower class. This is especially telling since even though the lower class only accounts for around 30% of the population (Pew Research Center, 2015), they still have the highest number of individuals killed by the police.

### Graph 2

```{r racial_breakdown, message=FALSE}
ggplot(data = police_killings, aes(state))+
  coord_flip()+
  geom_bar(aes(x = state, fill = raceethnicity))+
    labs(
    title = "Racial Breakdown of People Killed by the Police in 2015", 
    caption = "Source: fivethirtyeight"
  )+
  scale_x_discrete(name = "State Abbreviation")+
  scale_y_continuous(name = "Number of Individuals Killed")
```



## Word count

```{r word_count, message=FALSE, echo=FALSE}
text_stats()
```

## Standards

In this assignment, we attempted the following [standards](https://beanumber.github.io/sds192/standards.html):

- `r standard_button("markdown")`: We met the Markdown standard because we...
- `r standard_button("aesthetics")`: We met the Aesthetics standard because we...


