---
title: "SDS 192 Assignment"
subtitle: "SDS 192"
author: "Malaika Kironde, Jack Kendrick, Lika Mikhelashvili"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE
)
library(tidyverse)
library(sds192)
library(fivethirtyeight)
library(ggthemes)
```

## Introduction

```{r}
# Adding in column for Region to dataset
state_region <- data.frame(state = state.abb, region = state.region)
state_region <- rbind(state_region, data.frame(state = "DC", region = "South"))

police_data <- merge(police_killings, state_region, by = "state")
police_data
```

### Graph 1

```{r graph1, message=FALSE}
#Plot Code - Wrapping by Race
ggplot(data=police_data, aes(h_income)) +
  geom_histogram() +
  theme_fivethirtyeight()+
  theme(axis.title = element_text())+
  scale_fill_fivethirtyeight()+
  facet_wrap(~raceethnicity) + 
  scale_x_continuous(name = "Tract-level Median Household Income (h_income)")+
  scale_y_continuous(name = "Number of Individuals Killed")+
  labs(
    title = "Median Household Income Distribution Among People Killed by the Police in 2015", 
    subtitle = "Distribution by Race",
    caption = "Source: fivethirtyeight"
  )
```

### Graph 2

The graphs below show the racial breakdowns of the individuals killed by the police in 2015 in each state (Figure 2.1) and region (Figure 2.2). Each race is mapped to a specific color. In figure 2.1 it is evident that in most states, the majority of the killings were of White race. However, a large proportion of Black individuals were killed in almost all states. Such disproportional trend is more evident in figure 2.2, where among all minority groups (Asian/Pacific Islander, Black Hispanic/Latino, Native American), a drastically high number of Black individuals are killed in almost all regions. 

```{r state_breakdown, message=FALSE}
ggplot(data = police_data, aes(state))+
  coord_flip()+
  geom_bar(aes(x = state, fill = raceethnicity))+
    labs(
    title = "Figure 2.1: Racial Breakdown of Police Killings in 2015", 
    subtitle = "States",
    caption = "Source: fivethirtyeight",
    fill = "Race/Ethnicity"
  )+
  scale_x_discrete(name = "State Abbreviation")+
  scale_y_continuous(name = "Number of Individuals Killed")
```

```{r region_breakdown, message= FALSE}
ggplot(data = police_data, aes(region))+
  geom_bar(aes(x = region, fill = raceethnicity))+
  labs(title = "Figure 2.2: Racial Breakdown of Police Killings in 2015",
  subtitle = "US Regions",
  caption = "Source: fivethirtyeight",
  fill = "Race/Ethnicity"
  )+
  scale_x_discrete(name = "Region")+
  scale_y_continuous(name = "Number of Individuals Killed")
```


## Word count

```{r word_count, message=FALSE, echo=FALSE}
text_stats()
```

## Standards

In this assignment, we attempted the following [standards](https://beanumber.github.io/sds192/standards.html):

- `r standard_button("markdown")`: We met the Markdown standard because we...
- `r standard_button("aesthetics")`: We met the Aesthetics standard because we...


